---
title: "Assignment 4"
author: "79546 - Stephen K. Ng'etich"
output:
  pdf_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

```{r echo=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=6) 

```

# Pre-requisite
## Load Packages
```{r message = FALSE,warning = FALSE}
# Clear variables
rm(list=ls())

library(splines)
library(npreg)
library(ISLR)
library(dplyr)
library(ggplot2)
library(gam)
```
## Load Dataset

```{r}
summary(Wage)
dataset = Wage
```

# Exploratory Data Analysis

```{r}
ggplot(Wage, aes(x = education, y = wage, color = health_ins)) +  # ggplot function
  geom_boxplot()
```

# Model Selection  

```{r}
lm_model =  lm(wage ~ year+age ,data = dataset)
summary(lm_model)
```
Generate a plot of the wages over age
```{r}
plot(Wage$age,Wage$wage)
```
Get min/max values of age using the range() function

```{r}
agelims = Wage %>%
    select(age) %>%
    range

agelims
```
Generate a sequence of age values spanning the range

```{r}
# Generate a sequence of age values spanning the range
age_grid = seq(from = min(agelims), to = max(agelims))
age_grid
```
## Regression Splines    
```{r}
# Fit a regression spline using basis functions
fit =  lm(wage~ns(age,df=6), data=Wage)

# Predict the value of the generated ages, 
# returning the standard error using se = TRUE
pred = predict(fit, newdata = list(age = age_grid), se = TRUE)

# Compute error bands (2*SE)
se_bands = with(pred, cbind("upper" = fit+2*se.fit, 
                            "lower" = fit-2*se.fit))

# Plot the spline and error bands
ggplot() +
  geom_point(data = Wage, aes(x = age, y = wage),colour="gray74") +
  geom_line(aes(x = age_grid, y = pred$fit), color = "#0000FF") + 
  geom_ribbon(aes(x = age_grid, 
                  ymin = se_bands[,"lower"], 
                  ymax = se_bands[,"upper"]), 
              alpha = 0.3) +
  xlim(agelims)


```
  
## Smoothing Splines  
  
```{r ,warning=FALSE}
# Fit 2 smoothing splines
fit_smooth = with(Wage, smooth.spline(age, wage, df = 16))
fit_smooth_cv = with(Wage, smooth.spline(age, wage, cv = TRUE))


# Plot the smoothing splines
ggplot() +
  ggtitle("Smoothing Splines") +
  xlab("Age")+
  ylab("Wage") +
  geom_point(data = Wage, aes(x = age, y = wage),colour="gray74") +
  geom_line(aes(x = fit_smooth$x, y = fit_smooth$y, 
                color = "16 degrees of freedom"))  +
  geom_line(aes(x = fit_smooth_cv$x, y = fit_smooth_cv$y, 
                color = "6.8 effective degrees of freedom")) +
  theme(legend.position = 'bottom')+ 
  labs(title = "Smoothing Splines", colour="")
```
  
## Natural Splines  
```{r}
natural.model = lm(wage~ns(age,df=4),data = Wage)
ns_predict = predict(natural.model, newdata= list(age = age_grid), se=TRUE)


# Compute error bands (2*SE)
se_bands = with(ns_predict, cbind("upper" = fit+2*se.fit, 
                            "lower" = fit-2*se.fit))

# Plot the spline and error bands
ggplot() +
  ggtitle("Natural Splines") +
  xlab("Age")+
  ylab("Wage") +
  geom_point(data = Wage, aes(x = age, y = wage),colour="gray74") +
  geom_line(aes(x = age_grid, y = ns_predict$fit), color = "red3") + 
  geom_ribbon(aes(x = age_grid, 
                  ymin = se_bands[,"lower"], 
                  ymax = se_bands[,"upper"]), 
              alpha = 0.3) +
  xlim(agelims)


```

## Using GAM Function  

```{r}
gam1<-gam(wage~s(age,df=6)+s(year,df=6)+education ,data = Wage)
par(mfrow=c(1,3)) #to partition the Plotting Window
summary(gam1)

#Plotting the Model
plot(gam1,se = TRUE,col="green") 
#se stands for standard error Bands
```
From the above plots:  
1. wages increases with age then decreases at around 60.  
2. There is decrease on salary at around year 2007 or 2008.  
3. Wages increases with the level of education  

## Logistic Regression Using GAM  
We can have the logistic model to predict a person can earn more than or less than 250 based on the age,year and education.

```{r}
logitgam1<-gam(I(wage > 250) ~ s(age,df=4) + s(year,df=4) + education ,data=Wage,family=binomial)
par(mfrow=c(1,3))
plot(logitgam1,se=T,col="green")
```

