---
title: "Assignment 3"
author: "79546 - Stephen K. Ng'etich"
output:
  pdf_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

```{r echo=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=6) 

```

# Pre-requisite
## Load packages
```{r message = FALSE,warning = FALSE}
# Clear variables
rm(list=ls())

library(readxl)
library(dplyr)
library(tidyverse)
library(lattice)
library(leaps)
library(MASS)

```
## Load dataset

```{r}

# Load Dataset
dataset <- read_excel("dataset/Sales_Dataset.xlsx")

```

# Exploratory Data Analysis

```{r}
summary(dataset)
```
```{r}
plot(dataset)
```
From the visual inspection of the plot, the variables are all correlated since the plot shows a linear relationship and heads on one line.

# Model Selection

## TV and polynomials

```{r}

tv_model = lm(Sales ~poly(TV,degree=2,raw = TRUE), data=dataset)  

summary(tv_model)

```
Polynomial of TV is not significant


## Newspaper and polynomials

```{r}

np_model = lm(Sales ~poly(Newspaper,degree=2,raw = TRUE), data=dataset)  
summary(np_model)

```
Polynomial of newspaper is significant in predicting sales

## Radio and polynomials

```{r}

radio_model = lm(Sales ~poly(Radio,degree=2,raw = TRUE), data=dataset)  

summary(radio_model)

```
polynomial of Radio is not significant since it has a value greater than 0.05

## BillBoard and its polynomials

```{r}

bb_model = lm(Sales ~poly(BillBoard,degree=2,raw = TRUE), data=dataset)  

summary(bb_model)

```
Polynomial of billboard is not statistically significant.

## Selecting the best predictors of Sales.

```{r}
lm_model = lm(Sales ~ . , data=dataset)
summary(lm_model)
```
Radio, Newspaper and TV has p-value greater than 0.05.This mean that they are not statically significant in predicting Sales.However, BillBoard is the best predictor for predicting sales since its p-values is less than 0.05.The result model equation can be written as $$Y= 27.125 + 9.0836 \times BillBoard$$

## Interaction of Billboard and Newspaper

```{r}
lm_model1 = lm(Sales ~ BillBoard*Newspaper, data=dataset)
summary(lm_model1)
```
The interaction between newspaper and Billboard is not significant

## Interaction of Billboard and TV

```{r}
lm_model2 = lm(Sales ~ BillBoard*TV, data=dataset)
summary(lm_model2)
```
The interaction between newspaper and Billboard is not significant

## Interaction of Billboard,Newspaper,TV and Radio

```{r}
lm_model2 = lm(Sales ~ BillBoard +Newspaper+BillBoard*TV*Radio*Newspaper, data=dataset)
summary(lm_model2)
```

```{r}
lm_model2 = lm(Sales ~ TV*Radio*Newspaper, data=dataset)
summary(lm_model2)
```
Omitting BillBoard,interaction of TV:Radio and TV:Newspaper are significant in predicting sales

```{r}
lm_model2 = lm(Sales ~ BillBoard + TV + Radio +Newspaper + TV*Radio, data=dataset)
summary(lm_model2)
```