---
title: "Assignment 2"
author: "79546 - Stephen K. Ng'etich"
output:
  pdf_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

```{r echo=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=6) 

```

# Pre-requisite
## Load packages
```{r message = FALSE,warning = FALSE}
# Clear variables
rm(list=ls())

library(readxl)
library(dplyr)
library(tidyverse)
library(lattice)
library(leaps)
library(MASS)

```
## Load dataset

```{r}

# Load Dataset
dataset <- read_excel("dataset/Dataset2.xlsx")

```

# Exploratory Data Analysis

Convert `Region` column data type to `factor` 
```{r}
dataset$Region <- as.factor(dataset$Region)
```




## Summary statistic

```{r}
summary(dataset)
```

```{r}

plot(dataset)
```

# Model Selection
## Build the linear model  

```{r}
power_lm_model = lm(Power~ .,data=dataset)
summary(power_lm_model)
```
From the summary, Homes show coefficient of Rates are `NA` this means that it does not add any information to the model.

```{r}
# Remove Rates from the predictor variables
dataset1 = subset(dataset,select = -c(Rates))

power_lm_model =  lm(Power~ ., data=dataset1)
summary(power_lm_model)
```

The following stepwise regression methods are used for model selection  
1. Forward  
2. Backward  
3. Both

## Forward Selection  
```{r}
modelfoward =  stepAIC(power_lm_model,direction="forward",trace = TRUE)
step(modelfoward)
```
From forward selection of the optimal model is $$Power = Homes + Sequence$$ with the lowest *AIC of  -152.63* 

## Backward Selection 
```{r}
modelfoward =  stepAIC(power_lm_model,direction="backward",trace = TRUE)
```

From backward selection,the optimal model is $$Power = Homes + Sequence$$ with the lowest *AIC of  -152.63* 


```{r}
modelfoward =  stepAIC(power_lm_model,direction="both",trace = TRUE)
step(modelfoward)
```
From both forward and backward selection,the optimal model is $$Power = 1.8676Homes + 0.5864Sequence - 2.6955$$ with the lowest *AIC of  -152.63* 

