---
title: "Assignment 1"
author: "79546 - Stephen K. Ng'etich"
output:
  pdf_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

```{r echo=FALSE}
knitr::opts_chunk$set(fig.width=5, fig.height=5) 

```

# Pre-requisite

## Install package

```{r message = FALSE,warning = FALSE}
#install.packages("readxl")
#install.packages('dplyr')
```

## Load package

```{r message = FALSE,warning = FALSE}
# Clear variables
rm(list=ls())

library(readxl)
library(dplyr)

```

## Load dataset

```{r}

# Load Dataset
dataset <- read_excel("dataset/Data1.xlsx")

salaries = dataset$Salaries
experience = dataset$`Years of Experience`
age = dataset$Age
gender = dataset$Gender
region = dataset$Region

```

# Exploratory Data Analysis

View the summary statistics of each numeric column

```{r}
summary(dataset)
```

## Salaries Distribution

```{r}
boxplot(salaries)
hist(salaries)
```

From the histogram and the box plots, the data show the following:\
1. The salaries data is continuous\
2. A lot of people are earning 3000 to 3200.\
3. Salaries is negative skewed because the mean is less than the median

## Age Distribution

```{R}
hist(age)
boxplot(age)
```

Age is negatively skewed with people 30 to 32 highly represented.

## Experience Distribution

```{r}
hist(experience)
```

This data represent people have more than `10 years experience`

## Age Distribution by Gender

```{r}
boxplot(age~gender,
        main="Age Distribution by Gender")
```

```{r}
boxplot(salaries~gender,
        main="Salary distribution by Gender")
```

## Salary Distribution per Region

```{r}

boxplot(salaries~region, main="Salary Disribution per Region")
```

Region 2 has the highest salary variation

## Correlation among the variables

```{R}

plot(data.frame(salaries,age,region,experience))
```

Salaries, age and experience are correlated this means that experience is a reasonable predictor of salaries

# Model Estimation

```{r}
lm1 =  lm(formula= salaries~experience+age)
summary(lm1)
```

**Years of Experience** is a good predictor of salaries since the p-value is less than 0.05

\$\$Salaries = 1679.855 , + , 91`Years of experience`\$\$
