---
title: "R Notebook"
output: pdf_document

  
---
```{r}
knitr::opts_chunk$set(fig.width=6, fig.height=6) 

```

# Assignment 1

## Pre-requisite
### install package
```{r message = FALSE,warning = FALSE}
#install.packages("readxl")
#install.packages('dplyr')
```

### load package
```{r}
# Clear variables
rm(list=ls())

library(readxl)
library(dplyr)

```
### Load dataset

```{r}

# Load Dataset
dataset <- read_excel("dataset/Data1.xlsx")

salaries = dataset$Salaries
experience = dataset$`Years of Experience`
age = dataset$Age
gender = dataset$Gender
region = dataset$Region

```

## Exploratory Data Analysis
View the summary of each numeric column

```{r}
summary(dataset)
```
### Salaries
```{r}
summary(salaries)
boxplot(salaries)
hist(salaries)
```
From the histogram and the box plots, the data show the following:  
1. The salaries data is continuous  
2. A lot of people are earning 3000 to 3200.  
3. Salaries is negative skewed because the mean is less than the median

### Age
```{R}
hist(age)
boxplot(age)
```

Age is negatively skewed with people 30 to 32 highly represented.

```{r}
hist(experience)
```

This data represent people have more than 10 years experience

```{r}
boxplot(age~gender,
        main="Age distribution by Gender")
```
```{r}
boxplot(salaries~gender,
        main="Age distribution by Gender")
```

```{r}

boxplot(salaries~region, main="Salary Disribution per Region")
```
Region 2 has the highest salary variation


```{R}

plot(data.frame(salaries,age,region,experience))
```
Salaries and experience are correlated this means that experience is a reasonable predictor of size

```{r}
simple.regression = lm(formula= salaries~experience)
summary(simple.regression)
```
```{r}
lm1 =  lm(dataset,formula = Salaries~.)
summary(lm1)
```

